# Data transformation

Data cleaning is one of the major tasks before doing the data exploration.

The data we used had some missing values but they were read as empty strings. Doing null value analysis on empty strings is difficult, so the empty strings were converted into NA for all the columns.

The data had some misprints in the age for perpetrators. Ages such as 1020, 940, 224 which cannot be possible in reality were present. Such values were removed from the dataset.

We also created some additional columns of day of week, year, month, day so as to directly use these columns in data exploration.
The interactive graph generated required data to be in a different form and also multiple files were generated for it which can be seen in the data folder

```{r}
urlfile<-'https://raw.githubusercontent.com/EDAVGroup29/NYC_ShootingAnalysis/main/data/NYPD_Shooting_Incident_Data_Historic.csv'
shooting<-read.csv(urlfile)
```

```{r}
library(naniar)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)
```


```{r}
shooting[shooting==""]<-NA
shooting$OCCUR_DATE <- as.POSIXct(shooting$OCCUR_DATE, format = "%m/%d/%Y")
shooting$year <- format(as.Date(shooting$OCCUR_DATE), "%Y")
shooting$month <- format(as.Date(shooting$OCCUR_DATE), "%m")
shooting$day <- format(as.Date(shooting$OCCUR_DATE), "%d")
shooting$weekday <- weekdays(as.Date(shooting$OCCUR_DATE))
shooting = shooting[order(shooting$OCCUR_DATE),]

```


```{r eval = FALSE}
#Transforming data for interactive chart
month_boro<-shooting %>%
  dplyr::group_by(month,BORO,year)%>%
  dplyr::summarise(num_shootings = n_distinct(INCIDENT_KEY))

month_boro <- month_boro %>% spread(year, num_shootings)
intereactive_data = month_boro[month_boro$BORO == "BRONX",]
write.csv(intereactive_data, "Bronx_data.csv")
intereactive_data = month_boro[month_boro$BORO == "BROOKLYN",]
write.csv(intereactive_data, "Brooklyn_data.csv")
intereactive_data = month_boro[month_boro$BORO == "MANHATTAN",]
write.csv(intereactive_data, "Manhattan_data.csv")
intereactive_data = month_boro[month_boro$BORO == "QUEENS",]
write.csv(intereactive_data, "Queens_data.csv")
intereactive_data = month_boro[month_boro$BORO == "STATEN ISLAND",]
write.csv(intereactive_data, "Staten_data.csv")

```

